Vegetation regrowth/encroachment (gjengroing)

### 2023-08-10 DRAFT_gjengroing.Rmd
Update DRAFT_gjengroing.Rmd
exchanged path files to P: and R:

```{r}
ssb10km <- st_read('/data/R/GeoSpatialData/Population_demography/Norway_SSB/Original/ssb_10km/ssb10km.shp')%>%
  # mutate the id value so that it aligns with the data coming from GEE
  mutate(SSBID = as.numeric(SSBID)/1000)
# Transform to the correct projection
ssb10km <- st_transform(ssb10km, st_crs(regions))

ssb50km <- st_read('/data/R/GeoSpatialData/Population_demography/Norway_SSB/Original/ssb_50km/ssb50km.shp')%>%
  # mutate the id value so that it aligns with the data coming from GEE
  mutate(SSBID = as.numeric(SSBID)/1000)
# Transform to the correct projection
ssb50km <- st_transform(ssb50km, st_crs(regions))
```

@@

```{r}
bioclim <- st_read('/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp') %>%
  mutate(NAVN = ifelse(NAVN == 'S°rboreal', 'Sørboreal', NAVN))
bioclim <- st_transform(bioclim, st_crs(regions))
```

#### NiN polygons
[NiN polygons](https://kartkatalog.miljodirektoratet.no/dataset/Details/2050) are imported here from the NINA servers:
```{r}
nin <- st_read('/data/R/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/versjon20221231/Natur_Naturtyper_NiN_norge_med_svalbard_25833/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb')%>%
  # Simplify ecosystem names
  mutate(hovedøkosystem = recode(hovedøkosystem, 
                                 "Våtmark" = 'Våtmark',
                                 "Skog" = "Skog",
                                 "Ingen" = "Ingen",
                                 "Fjell" = "Fjell",
                                 "Semi-naturlig mark" = 'Semi-naturlig',
                                 "Naturlig åpne områder i lavlandet" = 'Naturlig åpne',
                                 "Naturlig åpne områder under skoggrensa" = 'Naturlig åpne')) %>% 
  # Fix any invalid geometries
  mutate(validGeo = st_is_valid(SHAPE)) %>%
  filter(validGeo) %>%
  # Filter out the ecosystems we are not interested in
  filter(!hovedøkosystem %in% c('Skog','Ingen','Fjell')) %>%
  # Get the main ecosystem codes in case we need them later
  mutate(hovedtype = substr(ninkartleggingsenheter, 4, 6),
         hovedtype = str_remove(hovedtype, '-'),
         id = identifikasjon_lokalid) %>%
  # Drop polygons with no condition score
  drop_na(tilstand) %>%
  dplyr::select(id, hovedøkosystem, hovedtype, naturtypekode, tilstand)
```

@@

```{r}
# Import data coverages per SSB 10km grid cell
areaCovers <- read_csv('./DATA/From_GEE/area_cover_grid.csv') %>%
  dplyr::select(-'.geo', -'system:index')

@@

elevStrata <- st_read('./DATA/From_GEE/elevation_stratified.shp') %>%
  mutate(elevBand = ifelse(mode == 1, '0-300',
                           ifelse(mode == 2, '300-600',
                                  ifelse(mode == 3, '600-900',
                                         ifelse(mode == 4, '900-1200', '>1200')))),
         elevBand = factor(elevBand, levels=c('0-300','300-600','600-900','900-1200','>1200')))
elevStrata <- st_transform(elevStrata, st_crs(regions))

@@

  # Export to file for reference
  refVaatmark %>%
    st_write('./DATA/refVaatmark.csv')
  refAapne %>%
    st_write('./DATA/refAapne.csv')

@@

  refVaatmark  <- read_csv('./DATA/refVaatmark.csv')
  refAapne <- read_csv('./DATA/refAapne.csv')

@@

  vaatmarkIndexPolySpat <- st_read('./OUTPUT/vaatmark_index.shp')%>%
    mutate(`system:index` = id,
           elevation = elev, 
           vegClimZoneLab = vegZn) %>%
    dplyr::select(-id, -elev, -vegZn)
  aapneIndexPolySpat <- st_read('./OUTPUT/aapne_index.shp')%>%
    mutate(`system:index` = id,
           elevation = elev, 
           vegClimZoneLab = vegZn) %>%
    dplyr::select(-id, -elev, -vegZn)

@@

  # Write out for import later
  vaatmarkIndexStrata %>% write_csv('./DATA/vaatmarkIndexStrata.csv')
  aapneIndexStrata %>% write_csv('./DATA/aapneIndexStrata.csv')

@@

  vaatmarkIndexStrata <- read_csv('./DATA/vaatmarkIndexStrata.csv')
  aapneIndexStrata <- read_csv('./DATA/aapneIndexStrata.csv')

@@

  vaatmarkIndexGrid <- st_read('./OUTPUT/vaatmark_index_grid.shp')
  aapneIndexGrid <- st_read('./OUTPUT/aapne_index_grid.shp')

@@

  vaatmarkIndexRegion <- st_read('./OUTPUT/vaatmark_index_region.shp')
  aapneIndexRegion <- st_read('./OUTPUT/aapne_index_region.shp')

@@

```{r}
include_graphics('./DATA/images/val_1.jpg')
```

@@

```{r}
include_graphics('./DATA/images/val_2.jpg')
```

@@

```{r}
include_graphics('./DATA/images/val_3.jpg')
```

@@

```{r}
include_graphics('./DATA/images/val_4.jpg')
```

@@

```{r}
include_graphics('./DATA/images/val_5.jpg')
```

@@

if (runFromScratch){
  # Polygon level
  vaatmarkIndexPolySpat  %>%
    #mutate(area = as.numeric(st_area(geometry)))%>% 
    #filter(area > 7500) %>%
    mutate(id = `system:index`,
           elev = elevation, 
           vegZn = vegClimZoneLab) %>%
    dplyr::select(-"system:index", -vegClimZoneLab, -elevation) %>%
    st_write('./OUTPUT/vaatmark_index.shp', append=FALSE)

  aapneIndexPolySpat %>% 
    #mutate(area = as.numeric(st_area(geometry)))%>% 
    #filter(area > 7500) %>%
    mutate(id = `system:index`,
           elev = elevation, 
           vegZn = vegClimZoneLab) %>%
    dplyr::select(-"system:index", -vegClimZoneLab, -elevation) %>%
    st_write('./OUTPUT/aapne_index.shp', append=FALSE)

  # Grid level
  vaatmarkIndexGrid %>%
    st_write('./OUTPUT/vaatmark_index_grid.shp', append = FALSE)
  aapneIndexGrid %>%
    st_write('./OUTPUT/aapne_index_grid.shp', append = FALSE)

  # Region level
  vaatmarkIndexRegion %>%
    st_write('./OUTPUT/vaatmark_index_region.shp', append = FALSE)
  aapneIndexRegion %>%
    st_write('./OUTPUT/aapne_index_region.shp', append = FALSE)
}

>>>>

```{r}
ssb10km <- st_read('R:/GeoSpatialData/Population_demography/Norway_SSB/Original/ssb_10km/ssb10km.shp')%>%
  # mutate the id value so that it aligns with the data coming from GEE
  mutate(SSBID = as.numeric(SSBID)/1000)
# Transform to the correct projection
ssb10km <- st_transform(ssb10km, st_crs(regions))

ssb50km <- st_read('R:/GeoSpatialData/Population_demography/Norway_SSB/Original/ssb_50km/ssb50km.shp')%>%
  # mutate the id value so that it aligns with the data coming from GEE
  mutate(SSBID = as.numeric(SSBID)/1000)
# Transform to the correct projection
ssb50km <- st_transform(ssb50km, st_crs(regions))
```

@@

```{r}
bioclim <- st_read('R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp') %>%
  mutate(NAVN = ifelse(NAVN == 'S°rboreal', 'Sørboreal', NAVN))
bioclim <- st_transform(bioclim, st_crs(regions))
```

#### NiN polygons
[NiN polygons](https://kartkatalog.miljodirektoratet.no/dataset/Details/2050) are imported here from the NINA servers:
```{r}
nin <- st_read('R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/versjon20221231/Natur_Naturtyper_NiN_norge_med_svalbard_25833/Natur_Naturtyper_NiN_norge_med_svalbard_25833.gdb')%>%
  # Simplify ecosystem names
  mutate(hovedøkosystem = recode(hovedøkosystem, 
                                 "Våtmark" = 'Våtmark',
                                 "Skog" = "Skog",
                                 "Ingen" = "Ingen",
                                 "Fjell" = "Fjell",
                                 "Semi-naturlig mark" = 'Semi-naturlig',
                                 "Naturlig åpne områder i lavlandet" = 'Naturlig åpne',
                                 "Naturlig åpne områder under skoggrensa" = 'Naturlig åpne')) %>% 
  # Fix any invalid geometries
  mutate(validGeo = st_is_valid(SHAPE)) %>%
  filter(validGeo) %>%
  # Filter out the ecosystems we are not interested in
  filter(!hovedøkosystem %in% c('Skog','Ingen','Fjell')) %>%
  # Get the main ecosystem codes in case we need them later
  mutate(hovedtype = substr(ninkartleggingsenheter, 4, 6),
         hovedtype = str_remove(hovedtype, '-'),
         id = identifikasjon_lokalid) %>%
  # Drop polygons with no condition score
  drop_na(tilstand) %>%
  dplyr::select(id, hovedøkosystem, hovedtype, naturtypekode, tilstand)
```

@@

```{r}
# Import data coverages per SSB 10km grid cell
areaCovers <- read_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/From_GEE/area_cover_grid.csv') %>%
  dplyr::select(-'.geo', -'system:index')

@@

elevStrata <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/From_GEE/elevation_stratified.shp') %>%
  mutate(elevBand = ifelse(mode == 1, '0-300',
                           ifelse(mode == 2, '300-600',
                                  ifelse(mode == 3, '600-900',
                                         ifelse(mode == 4, '900-1200', '>1200')))),
         elevBand = factor(elevBand, levels=c('0-300','300-600','600-900','900-1200','>1200')))
elevStrata <- st_transform(elevStrata, st_crs(regions))

@@

  # Export to file for reference
  refVaatmark %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/refVaatmark.csv')
  refAapne %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/refAapne.csv')

@@

  refVaatmark  <- read_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/refVaatmark.csv')
  refAapne <- read_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/refAapne.csv')

@@

  vaatmarkIndexPolySpat <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index.shp')%>%
    mutate(`system:index` = id,
           elevation = elev, 
           vegClimZoneLab = vegZn) %>%
    dplyr::select(-id, -elev, -vegZn)
  aapneIndexPolySpat <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index.shp')%>%
    mutate(`system:index` = id,
           elevation = elev, 
           vegClimZoneLab = vegZn) %>%
    dplyr::select(-id, -elev, -vegZn)

@@

  # Write out for import later
  vaatmarkIndexStrata %>% write_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/vaatmarkIndexStrata.csv')
  aapneIndexStrata %>% write_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/aapneIndexStrata.csv')

@@

  vaatmarkIndexStrata <- read_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/vaatmarkIndexStrata.csv')
  aapneIndexStrata <- read_csv('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/aapneIndexStrata.csv')

@@

  vaatmarkIndexGrid <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index_grid.shp')
  aapneIndexGrid <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index_grid.shp')

@@

  vaatmarkIndexRegion <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index_region.shp')
  aapneIndexRegion <- st_read('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index_region.shp')

@@

```{r}
include_graphics('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/images/val_1.jpg')
```

@@

```{r}
include_graphics('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/images/val_2.jpg')
```

@@

```{r}
include_graphics('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/images/val_3.jpg')
```

@@

```{r}
include_graphics('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/images/val_4.jpg')
```

@@

```{r}
include_graphics('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/DATA/images/val_5.jpg')
```

@@

if (runFromScratch){
  # Polygon level
  vaatmarkIndexPolySpat  %>%
    #mutate(area = as.numeric(st_area(geometry)))%>% 
    #filter(area > 7500) %>%
    mutate(id = `system:index`,
           elev = elevation, 
           vegZn = vegClimZoneLab) %>%
    dplyr::select(-"system:index", -vegClimZoneLab, -elevation) %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index.shp', append=FALSE)

  aapneIndexPolySpat %>% 
    #mutate(area = as.numeric(st_area(geometry)))%>% 
    #filter(area > 7500) %>%
    mutate(id = `system:index`,
           elev = elevation, 
           vegZn = vegClimZoneLab) %>%
    dplyr::select(-"system:index", -vegClimZoneLab, -elevation) %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index.shp', append=FALSE)

  # Grid level
  vaatmarkIndexGrid %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index_grid.shp', append = FALSE)
  aapneIndexGrid %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index_grid.shp', append = FALSE)

  # Region level
  vaatmarkIndexRegion %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/vaatmark_index_region.shp', append = FALSE)
  aapneIndexRegion %>%
    st_write('P:/41201785_okologisk_tilstand_2022_2023/gjengroing/R_project/OUTPUT/aapne_index_region.shp', append = FALSE)
}
